### 2017-05-23 sheets/basis-3/sheet-basis-3.tex
Sample solution of Exercise Basis 3.4
Add exercise 6 solution sketch

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like 'God-given' or
  'defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

@@

@@

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F: \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && \forall n \in \NN\\
        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{ matrices } A
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f: K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is surjective on the objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
\end{answer}

>>>>

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like `God-given' or
  `defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by `choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

@@

\begin{answer}
  \begin{itemize}
    \item[(a)]
      $\Sym(f) = \lambda g.f \circ g$, where $g$ is a permutation\\
      $\Ord(f) = \lambda (a, b).(f(a), f(b))$, where $a, b \in X$
    \item[(b)]
      Let's consider a non-identity permutation $h$. We can build an ordering
      from it by taking $(x, h(x))$ as an element of ordering relation for
      $\forall x \in \text{ some finite set } X$, then filling the gaps by
      reflexivity and transitivity.

      However, it's not the case for an identity permutation $i$, as for that
      permutation the $i(x) = x$ which leaves us only with reflexivity pairs if
      we try to build an ordering from that bijection. Thus, we cannot define
      formation of an ordering from an identity permutation and therefore
      there's no natural transformation from $\Sym$ to $\Ord$.
    \item[(c)]
      $|\Sym(X)| = n$!\\
      $|\Ord(X)| = n$!
  \end{itemize}
  We have seen that there's no natural transformation $\Sym \to \Ord$.
  Nonetheless, we have also seen that they have the same number of
  elements and an isomorphism between them can actually be formed: it's not
  hard to construct all permutations or orderings of some finite set knowing
  its elements.

  This leads us to the conclusion that $\Sym(X) \iso \Ord(X)$ for all $X \in
  \cat{B}$, but not \emph{naturally} in $X \in \cat{B}$.
\end{answer}

@@

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F: \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && \forall n \in \NN\\
        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{matrices} A
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f: K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is surjective on the objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
\end{answer}
